---
title: "Metaflow S3 Client"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Metaflow S3 Client}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

The Metaflow S3 Client provides a convenient interface for interacting with Amazon S3 storage. This vignette focuses on using the S3 client to list objects in S3 buckets.

## Setting up the S3 Client

First, let's create an S3 client:

```{r, eval=FALSE}
library(metaflow)

# Create a new S3 client
s3 <- create_s3_client(
  tmproot = ".",
  bucket = "my-bucket",
  prefix = "my-prefix"
)

# Connect to S3
s3$connect()
```

The `create_s3_client` function initializes a new S3 client. You can specify the temporary root directory, bucket name, and prefix. The `connect()` method establishes the connection to S3.

### Customizing S3 Client Configuration

```{r, eval=FALSE}
# Set AWS role
s3$set_role("arn:aws:iam::123456789012:role/my-role")

# Set session variables
s3$set_session_vars(list(var1 = "value1", var2 = "value2"))

# Set client parameters
s3$set_client_params(list(param1 = "value1", param2 = "value2"))
```


## Listing Objects with list_paths

The `list_paths` method allows you to list objects at a specific S3 path. It returns a list of `S3Object` representations.

```{r, eval=FALSE}
# List objects in the root of the S3 path
root_objects <- s3$list_paths()

# List objects with specific keys
specific_objects <- s3$list_paths(c("path1", "path2"))

# Print the results
print(root_objects)
print(specific_objects)
```

## Recursive Listing with list_recursive

The `list_recursive` method allows you to list all objects under given prefixes, including those in subfolders.

```{r, eval=FALSE}
# List all objects recursively
all_objects <- s3$list_recursive()

# List objects recursively from specific prefixes
specific_recursive <- s3$list_recursive(c("prefix1", "prefix2"))

# Print the results
print(all_objects)
print(specific_recursive)
```

## Working with S3Objects

Both `list_paths` and `list_recursive` return lists of `S3Object` instances. Each `S3Object` has the following properties:

- `exists`: Logical indicating if the object exists in S3
- `url`: The S3 URL of the object
- `prefix`: The prefix requested that matches this object
- `key`: The key corresponding to this object
- `size`: The size of the object in bytes
- `has_info`: Logical indicating if the object contains additional metadata
- `metadata`: A list of user-defined metadata (if any)
- `content_type`: The content type of the S3 object
- `last_modified`: The last modified timestamp
- `encryption`: The server-side encryption type (if set)

You can access these properties for each object using `purrr` functions:

```{r, eval=FALSE}
library(purrr)

# Print key, size, last_modified, and URL for each object
root_objects %>%
  walk(~ cat(
    "Key:", .$key, "\n",
    "Size:", .$size, "bytes\n",
    "Last Modified:", .$last_modified, "\n",
    "URL:", .$url, "\n\n"
  ))

# Extract specific information from all objects
keys <- map_chr(root_objects, "key")
sizes <- map_dbl(root_objects, "size")
last_modified_times <- map_dbl(root_objects, "last_modified")

# Create a data frame of object information
object_info <- root_objects %>%
  map_dfr(~ list(
    key = .$key,
    size = .$size,
    last_modified = .$last_modified,
    url = .$url
  ))

print(object_info)
```

